\chapter{IMPLEMENTATION DETAILS}
Note: This begining of this chapter should be mid-point of your entire rport content. 
The implementation detials and result part should approximately be 50-60\% of total content.
Describe how design was translated to working system with sufficient detail for replication.
Some relevant sections might be given below. You can heavilty customize based on your project type. 


\section{Hardware/Software Module Implementation}
HARDWARE: Component selection with part numbers, interface circuitry, PCB design (tools, layout).
ML: Data collection/cleaning, feature engineering (scaling, encoding, augmentation), dataset statistics.
SOFTWARE: Major components/features, database schema, API implementations, UI components.
Include code snippets, diagrams, and justifications.

\section{Firmware/Model Implementation}
HARDWARE: Embedded software architecture (control loops, interrupts, RTOS), peripheral drivers.
ML: Frameworks used, model architecture code, hyperparameter tuning, training procedures (resources, schedules).
Edge ML: Model optimization (quantization, pruning), conversion (TFLite, ONNX), profiling results.
Include pseudocode/snippets for critical algorithms.

\section{Integration Process}
How subsystems connected, interface testing, communication protocol implementations.
Calibration procedures (sensors, controllers), system-level validation.
Include integration diagrams and debugging approaches.

\section{Testing and Validation}
Unit tests (components), integration tests (subsystems), system tests (end-to-end).
Stress testing (environmental, long-duration, boundaries), field testing.
ML: Evaluation results (confusion matrices, error distributions, baselines, ablation studies).
Include test coverage reports and performance charts.

\section{Challenges and Solutions}
Technical problems encountered and solutions developed. Demonstrates engineering problem-solving.
HARDWARE: Failures, noise, timing, power issues.
ML: Training challenges (gradients, convergence), deployment issues.
SOFTWARE: Bugs, bottlenecks, concurrency, scaling.
Include debugging screenshots, before/after comparisons.
